Notes on Control Plane Inference

### 2021-05-24 inference.tex
sketch isntrumentation equivalence proof

% sets
\newcommand{\Value}{\mathsf{Value}}
\newcommand{\BVExpr}{\mathsf{BVExpr}}
\newcommand{\BExpr}{\mathsf{BExpr}}
\newcommand{\Cmd}{\mathsf{Cmd}}
\newcommand{\Action}{\mathsf{Action}}
\newcommand{\Instr}{\mathsf{Instr}}
\newcommand{\Prog}{\mathsf{Prog}}
\newcommand{\GCL}{\mathsf{GCL}}
\newcommand{\Pkt}{\mathsf{Pkt}}
\newcommand{\Model}{\mathsf{Model}}
\newcommand{\Hdr}{\mathsf{Hdr}}
\newcommand{\Field}{\mathsf{Field}}
\newcommand{\Table}{\mathsf{Table}}
\newcommand{\Row}{\mathsf{Row}}
\newcommand{\List}{\mathsf{List}}
\newcommand{\Var}{\mathsf{Var}}
\newcommand{\SymbRow}{\mathsf{SymbRow}}
\newcommand{\State}{\mathsf{State}}

% record fields
\newcommand{\matches}{\mathsf{matches}}
\newcommand{\action}{\mathsf{action}}
\newcommand{\actions}{\mathsf{actions}}
\newcommand{\keys}{\mathsf{keys}}
\newcommand{\data}{\mathsf{data}}
\newcommand{\params}{\mathsf{params}}
\newcommand{\id}{\mathsf{id}}
\newcommand{\reach}{\mathit{\mathit{reach}}}
\newcommand{\hit}{\mathit{hit}}
\newcommand{\reads}{\mathit{reads}}

% commands
\newcommand{\assert}{\mathop{\mathsf{assert}}}
\newcommand{\assume}{\mathop{\mathsf{assume}}}
\newcommand{\apply}{\mathsf{apply}}
\newcommand{\choiceop}{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}
\newcommand{\choice}{\mathbin{\choiceop}}
\DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
\newcommand{\havoc}[1]{\mathop{\textsf{havoc}}#1}
\newcommand{\SKIP}{\mathsf{skip}}

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\nondet}{\mathsf{nondet}}
\newcommand{\ghostly}{\mathsf{ghostly}}
\newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}
\newcommand{\instr}{[?]}
\newcommand{\ginstr}{[\mathghost]}
\newcommand{\qe}{\textsc{Qe}}
\newcommand{\dom}{\textrm{dom}}

@@

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} \\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} \\
      & \mid & x & \textit{Variable}\\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation}\\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table}\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}\\
      & \mid & b \Rightarrow b & \textit{Implication}\\
    & \mid & e = e & \textit{Bitvector Equality}\\
    %% & \mid & \rho \in t & \textit{Table membership}\\
    %% & \mid & \forall \rho.~b & \textit{Row Universality}\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment}\\
      %% & \mid & \havoc \rho & \textit{Row Havoc} & (\iota) \\
      & \mid & \assert b & \textit{Assertion}\\
      & \mid & \assume b & \textit{Assumption} \\
      & \mid & t.\apply() & \textit{Table Application} \\
      & \mid & c;c & \textit{Sequential Composition} \\
      & \mid & c \choice c & \textit{Nondeterministic Choice} \\
      %% & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$.}
\label{fig:grammar}
\end{figure}

@@

\paragraph{Tables}
A table $t \in \Table$ is a record with a few fields described below:
\begin{enumerate}[align=left]
  \item[($t.\id$)] A table's unique identifier. A well-formed program can only have
    one occurence of each table identifier.
  \item[($t.\keys$)] A list of header field accesses indicating the match keys.
  \item[($t.\actions$)] A list of actions that can be executed. An action is a lambda
    expression $\lambda \vec x. c$ where the variables $x$ are in scope for the
    commands $c$, which $c$ are straight-line code only (see below).
\end{enumerate}
For now we assume that the default action is always specified by the controller.
That is, programs never `miss' in tables. See the paragraph on state below. We use $|l|$ to indicate the number of
elements in a list or vector $l$ such as $t.\keys$ or $t.\actions$.

@@

\begin{definition}[Total State]
  \label{def:total-state}
  A state $\sigma$ is total if for every $t \in \dom(\sigma)$, the following
  formula is valid:
  \[\bigwedge_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
\end{definition}

@@

@@

\paragraph{Command Evaluation}
Commands are evaluated similar to Avenir in Figure~\ref{fig:com-sem}. The
notable differences are the inclusion of assertions, the nondeterminism
operators, and a cleaner representation of table application.

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Pkt \to \mathcal{P}(\Pkt)_\error }\\
    \denote{h.f := e}^\sigma\pkt && \{\pkt\{h \mapsto \edenote{e}\pkt\}\} \\
    \denote{\assume b}^\sigma\pkt &&
    \begin{cases}
      \{\pkt\} & \pkt \models b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma\pkt &&
    \begin{cases}
      \{\pkt\} & \pkt \models b \\
      \error & \textit{otherwise}
    \end{cases} \\
    %% \denote{\havoc \rho}^\sigma\,\mu\,\pkt &&
    %% \{(\mu\{\rho \mapsto r\}, \pkt) \mid r \in \Row\} \\
    \denote{c_1;c_2}^\sigma\pkt&&
    \bigcup_{\pkt' \in \denote{c_1}^\sigma\pkt}\denote{c_2}^\sigma\pkt' \\
    \denote{c_1 \choice c_2}^\sigma\pkt &&
    \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt \\
    %% \denote{\{c\}}^\sigma\,\mu\,\pkt &&
    %% \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
    \denote{t.\apply()}^\sigma\,\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma\pkt,
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~r = \argmin_i\{\pkt \models b_i\} \\
      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
      \{\pkt\}, & \mathit{otherwise}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands}
  \label{fig:com-sem}
\end{figure}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  (h.f := e)[\sigma] && h.f := e   \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  %% (\{c\})[\sigma] && \{c[\sigma]\} \\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    (\\
    \quad \assume {r_1.\matches = t.\keys};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    )~\choice\\
    \phantom)~\,\vdots\\
    \phantom)\,\choice~(\\
    \quad \assume {r_n.\matches = t.\keys};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    )\choice (\\
    \quad \displaystyle \assume {\bigwedge_{i=1}^n r_n.\matches \neq t.\keys};\\
    \quad \SKIP \\
    )\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Substitution}
  \label{fig:table-elim}
\end{figure}

\begin{theorem}[Table Substitution]
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
  packet $\pkt \in \Pkt$, then for any state $\tau : \Value \to \Row^*$,
  \[\denote{c}^\sigma\pkt = \denote{c[\sigma]}^\tau\pkt\]
\end{theorem}
\begin{proof}
\todo[inline]{By induction on the structure of $c$.}
\end{proof}

@@

\begin{figure}[H]
  \[
  \begin{array}{r >{\triangleq}c l}
    \WP(h.f := e, b)
    && b[e/h.f] \\
    %% \WP(\havoc \rho, b)
    %% && \forall \rho.\,b \\
    \WP(\assert b_a, b)
    && b_a \wedge b \\
    \WP(\assume b_a, b)
    && b_a \Rightarrow b \\
    \WP(t.\apply,b) && \textrm{\textcolor{red}{undefined}} \\
    \WP(c_1;c_2, b)
    && \WP(c_1,\WP(c_2,b)) \\
    \WP(c_1 \choice c_2, b)
    && \WP(c_1,b) \wedge \WP(c_2,b)
  \end{array}
  \]
  \caption{Dijkstra's weakest precondition function}
  \label{fig:wp}
\end{figure}

@@

\begin{conjecture}
  For every $c \in \Cmd$, and $\sigma \in \State$, the following are equivalent:
  \begin{itemize}
  \item \(\state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\)
  \item \(\WP(c[\sigma], \TRUE)\)
  \end{itemize}
\end{conjecture}

@@

\section{Quantifier Elimination?}

>>>>

% sets
\newcommand{\Value}{\mathsf{Value}}
\newcommand{\BVExpr}{\mathsf{BVExpr}}
\newcommand{\BExpr}{\mathsf{BExpr}}
\newcommand{\Cmd}{\mathsf{Cmd}}
\newcommand{\Action}{\mathsf{Action}}
\newcommand{\Instr}{\mathsf{Instr}}
\newcommand{\Prog}{\mathsf{Prog}}
\newcommand{\GCL}{\mathsf{GCL}}
\newcommand{\Pkt}{\mathsf{Pkt}}
\newcommand{\Model}{\mathsf{Model}}
\newcommand{\Hdr}{\mathsf{Hdr}}
\newcommand{\Field}{\mathsf{Field}}
\newcommand{\Table}{\mathsf{Table}}
\newcommand{\Row}{\mathsf{Row}}
\newcommand{\List}{\mathsf{List}}
\newcommand{\Var}{\mathsf{Var}}
\newcommand{\SymbRow}{\mathsf{SymbRow}}
\newcommand{\State}{\mathsf{State}}
\newcommand{\Kind}{\mathsf{Kind}}

% record fields
\newcommand{\matches}{\mathsf{matches}}
\newcommand{\action}{\mathsf{action}}
\newcommand{\actions}{\mathsf{actions}}
\newcommand{\keys}{\mathsf{keys}}
\newcommand{\data}{\mathsf{data}}
\newcommand{\params}{\mathsf{params}}
\newcommand{\id}{\mathsf{id}}
\newcommand{\reach}{\mathit{\mathit{reach}}}
\newcommand{\hit}{\mathit{hit}}
\newcommand{\reads}{\mathit{reads}}
\newcommand{\hitAction}{\mathsf{hitAction}}
\newcommand{\miss}{\mathsf{miss}}

% commands
\newcommand{\assert}{\mathop{\mathsf{assert}}}
\newcommand{\assume}{\mathop{\mathsf{assume}}}
\newcommand{\apply}{\mathsf{apply}}
\newcommand{\choiceop}{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}
\newcommand{\choice}{\mathbin{\choiceop}}
\DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
\newcommand{\havoc}[1]{\mathop{\mathsf{havoc}}#1}
\newcommand{\SKIP}{\mathsf{skip}}

% matchkinds
\newcommand{\exact}{\ensuremath{\mathsf{exact}}}
\newcommand{\ternary}{\ensuremath{\mathsf{ternary}}}
\newcommand{\lpm}{\ensuremath{\mathsf{lpm}}}

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\nondet}{\mathsf{nondet}}
\newcommand{\ghostly}{\mathsf{ghostly}}
\newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}
\newcommand{\instr}{[?]}
\newcommand{\ginstr}{[\mathghost]}
\newcommand{\qe}{\textsc{Qe}}
\newcommand{\dom}{\textrm{dom}}
\newcommand{\fvs}{\textrm{fvs}}
\newcommand{\efvs}{\textrm{efvs}}
\newcommand{\bfvs}{\textrm{bfvs}}

%% operators
\newcommand{\bvand}{\mathop{\texttt{\&}}}
\newcommand{\match}{\mathrel{\texttt{{=}\raisebox{-2.5px}{\textasciitilde}}}}
\newcommand{\matchrow}{\mathsf{matchrow}}

@@

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} \\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} \\
      & \mid & x & \textit{Variable}\\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation}\\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} \\
    \multicolumn{3}{l}{\Kind} \\
    t & ::= & \exact  & \textit{Exact Match}\\
    & \mid & \ternary & \textit{Ternary Match} \\
    & \mid & \lpm & \textit{Longest Prefix Match} \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f : \kappa]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table}\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}\\
      & \mid & b \Rightarrow b & \textit{Implication}\\
    & \mid & e = e & \textit{Bitvector Equality}\\
    %% & \mid & \rho \in t & \textit{Table membership}\\
    %% & \mid & \forall \rho.~b & \textit{Row Universality}\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment} \\
      & \mid & \havoc \rho & \textit{Row Havoc} \\
      & \mid & \assert b & \textit{Assertion}\\
      & \mid & \assume b & \textit{Assumption} \\
      & \mid & t.\apply() & \textit{Table Application} \\
      & \mid & c;c & \textit{Sequential Composition} \\
      & \mid & c \choice c & \textit{Nondeterministic Choice} \\
      %% & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$.}
\label{fig:grammar}
\end{figure}

@@

\paragraph{Tables}
A table $t \in \Table$ is a record with a few fields described below:
\begin{enumerate}[align=left]
  \item[($t.\id$)] A table's unique identifier. A well-formed program can only have
    one occurence of each table identifier.
  \item[($t.\keys$)] A list of header field accesses indicating the match keys and their match kind.
  \item[($t.\actions$)] A list of actions that can be executed. An action is a lambda
    expression $\lambda \vec x. c$ where the variables $x$ are in scope for the
    commands $c$, which $c$ are straight-line code only (see below).
\end{enumerate}
For now we assume that the default action is always specified by the controller.
That is, programs never `miss' in tables. See the paragraph on state below. We use $|l|$ to indicate the number of
elements in a list or vector $l$ such as $t.\keys$ or $t.\actions$.

\paragraph{Match Kind}
There are three match kinds, \exact, \ternary, and \lpm. Exact matches (\exact)
indicate that the bits in the key must be the same as the match bits. Ternary
matches (\ternary) specify a match and a mask and the key must be equivalent up
to masking. Longest prefix matches (\lpm) specify a match and a length
$l$, the first $l$ bits of the key must be equivalent to the match.

@@

\begin{definition}[Total State]
  \label{def:total-state}
  A state $\sigma$ is total if for every $t \in \dom(\sigma)$, the following
  formula is valid:
  \[\bigvee_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
\end{definition}

@@

\todo[inline]{Can be Removed(?)}

@@

\paragraph{Command Evaluation}
Commands are evaluated similar to Avenir in Figure~\ref{fig:com-sem}. The
notable differences are the inclusion of assertions, the nondeterminism
operators, support for inexact matches, and a cleaner representation of table
application. The denotation function $\denote{-}$ is parameterized on a
$\sigma$, it is used in the definition of $t.\apply()$. If no $\sigma$ is
provided, the behavior of $t.\apply()$ is undefined.

In evaluating tables, we must condition on and evaluate matches. To do this, we
define an operator $t.\matchrow : \Row \to \BExpr$ which checks that the lists of
matches and keys are equivalent up to the corresponding match type.

First, define an auxiliary binary operator $(\match) : (\BVExpr \cup \BVExpr^2 \cup \BVExpr/\BVExpr) \to (\BVExpr \times \Kind) \to \BExpr$ that is $\TRUE$ iff the arguments are equivalent up
to the match kind $\kappa$. Formally we define this as follows:

\[
\begin{array}{r<{\!\!\!} l >{\triangleq}cl}
  e \match e' :& \exact && e = e' \\
  (e_v,e_m) \match e' :& \ternary && e_v \bvand e_m = e' \bvand e_m \\
  e/l \match  e' :& \lpm && e \bvand 1^l0^{|e| - l} = e' \bvand 1^l0^{|e| - l} \\
\end{array}
\]

Now we can define $t.\matchrow(r)$ as follows
\[t.\matchrow(r) \triangleq
   \bigwedge_{i = 0}^n r.\matches[i] \match t.\keys[i]
\]

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Pkt \to \mathcal{P}(\Pkt)_\error }\\
  \denote{x := e}^\sigma\pkt && \{\pkt\{x \mapsto \edenote{e}\pkt\}\} \\
  \denote{\havoc x}^\sigma\pkt && \{ \pkt\{x \mapsto v\} \mid v \in \Value \} \\
    \denote{\assume b}^\sigma\pkt &&
    \begin{cases}
      \{\pkt\} & \pkt \models b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma\pkt &&
    \begin{cases}
      \{\pkt\} & \pkt \models b \\
      \error & \textit{otherwise}
    \end{cases} \\
    %% \denote{\havoc \rho}^\sigma\,\mu\,\pkt &&
    %% \{(\mu\{\rho \mapsto r\}, \pkt) \mid r \in \Row\} \\
    \denote{c_1;c_2}^\sigma\pkt&&
    %% \bigcup_{\pkt' \in \denote{c_1}^\sigma\pkt}\denote{c_2}^\sigma\pkt' \\
    {\denote{c_2}^\sigma}^\dagger\left(\denote{c_1}^\sigma\pkt\right)\\
    \denote{c_1 \choice c_2}^\sigma\pkt &&
    \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt \\
    %% \denote{\{c\}}^\sigma\,\mu\,\pkt &&
    %% \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
    \denote{t.\apply()}^\sigma\,\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma\pkt,
      & \text{exists}~r = \argmin_i\{\pkt \models b_i\} \\
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~b_i = t.\matchrow(\sigma(t.\id)[i]) \\
      \{\pkt\}, & \textit{if no such } r \textit{ exists}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands. The dagger operator $(\cdot^\dagger) : (\Pkt \to
    \mathcal{P}(\Pkt)_\error) \to \mathcal{P}(\Pkt)_\error \to
    \mathcal{P}(\Pkt)_\error$ lifts a function that accepts packets to one that
    accepts elements of to the pointed powerset.}
  \label{fig:com-sem}
\end{figure}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  (x := e)[\sigma] && h.f := e \\
  (\havoc x)[\sigma] && \havoc x \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  %% (\{c\})[\sigma] && \{c[\sigma]\} \\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    (\\
    \quad \assume {t.\matchrow(r_1)};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    )~\choice\\
    \phantom)~\,\vdots\\
    \phantom)\,\choice~(\\
    \quad \assume {t.\matchrow(r_n)};\\
    \quad \assume {\bigwedge_{i=1}^{n-1}{\neg t.\matchrow(r_i)}};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    )\choice (\\
    \quad \assume {\bigwedge_{i=1}^n \neg t.\matchrow(r_i)};\\
    \quad \SKIP \\
    )\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Substitution}
  \label{fig:table-elim}
\end{figure}

\begin{theorem}[Table Substitution]
  \label{thm:subst}
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
  packet $\pkt \in \Pkt$,
  \[\denote{c}^\sigma\pkt = \denote{c[\sigma]}\,\pkt\]
\end{theorem}
\begin{proof}
   Proceed by induction on the structure of $c$.
  \begin{enumerate}[align=left]
    \item[$(c = (x := e))$] Immediate, since $(x:=e)[\sigma] = (x:=e)$.

    \item[$(c = \havoc x)$] Immediate, since $(\havoc x)[\sigma] = \havoc x$.

    \item[$(c = \assume b)$] Immediate, since $(\assume b)[\sigma] = (\assume b)$.

    \item[$(c = \assert b)$] Immediate, since $(\assert b)[\sigma] = (\assert b)$.

    \item[$(c = c_1;c_2)$] Let $\pkt,\sigma$ be arbitrary. Compute as follows
      \[\begin{array}{ >{\displaystyle}l l}
        \denote{(c_1;c_2)[\sigma]}\,\pkt \\
        = \denote{c_1[\sigma];c_2[\sigma]}\,\pkt  & \text{definition of } (\cdot)[\sigma]\\
        = \bigcup_{\pkt' \in \denote{c_1[\sigma]}\,\pkt} \denote{c_2[\sigma]}\pkt' & \text{definition of } \denote{\cdot}\\
        = \bigcup_{\pkt' \in \denote{c_1[\sigma]}\,\pkt} \denote{c_2}^\sigma\pkt' & \text{IH}(c_2) \\
        = \bigcup_{\pkt' \in \denote{c_1}^\sigma\pkt} \denote{c_2}^\sigma\pkt' & \text{IH}(c_1) \\
        = \denote{c_1;c_2}^\sigma\pkt & \text{definition of } \denote{\cdot}^\sigma
      \end{array}\]

    \item[$(c = c_1 \choice c_2)$] Let $\pkt, \sigma$ be arbitrary. Compute as follows:
      \[ \begin{array}{>{\displaystyle}ll}
          \denote{c_1\choice c_2[\sigma]}\,\pkt\\
          = \denote{c_1[\sigma] \choice c_2[\sigma]}\,\pkt & \text{definition of } (\cdot)[\sigma]\\
          = \denote{c_1[\sigma]}\,\pkt \cup \denote{c_2[\sigma]}\,\pkt & \text{definition of } \denote{\cdot}\\
          = \denote{c_1}^\sigma\pkt \cup \denote{c_2[\sigma]}\,\pkt & \text{IH}(c_1)\\
          = \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\,\pkt & \text{IH}(c_2)\\
          = \denote{c_1\choice c_2}^\sigma\pkt & \text{definition of } \denote{\cdot}^\sigma
          \end{array}
      \]

    \item[$(c = t.\apply())$] Let $\pkt,\sigma$ be arbitrary. Let $[r_1, \ldots,
      r_n] = \sigma(t)$.

      Compute as follows:
      \[
      \begin{array}{ll}
        \denote{(t.\apply())[\sigma]}\,\pkt\\
        = \denote{
          \begin{array}{l}
            \bigchoice_{i=1}^{n}
            \begin{array}{l}
              \assume t.\matchrow(r_i);\\
              \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
              t.\action[r_i.\action](r_i.data)
            \end{array} \\
            \choice \assume \bigwedge_{i=1}^{n} \neg  t.\matchrow(r_j); \SKIP
        \end{array}}\pkt \\
        = \begin{array}{l}
          \bigcup_{i=1}^{|\sigma(t)|}
          \denote{
            \begin{array}{l}
              \assume t.\matchrow(r_i);\\
              \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
              t.\action[\sigma(t)[i]](r_i.data)
            \end{array}}\,\pkt\\
            \cup \denote{\assume \bigwedge_{i=1}^{n} \neg  t.\matchrow(r_j); \SKIP}\,\pkt
        \end{array} \\
        \end{array}
      \]
      Now consider the two cases.

      \begin{enumerate}
      \item[\textit{Case 1}.] Assume that there is some $i \in \{1,\ldots,
        n\}$ such that $\pkt \models t.\matchrow(r_i)$. Let $i$ be the
        smallest such $i$. Then by definition $\denote{t.\apply()}^\sigma =
        \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. Since it is
        the smallest, it must be the that $\pkt \not\models t.\matchrow(r_j)$
        for $1 \leq j < i$.This is equivalent to saying that $\pkt
        \models\bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)$. Consequently,
        $\denote{t.\action[r_i.\action](r_i.\data)}\,\pkt \subseteq
        \denote{t.\apply()[\sigma]}$. In fact this is an equality.

        Because $i$ is the smallest index satisfying $\pkt \models
        t.\matchrow(r_i)$, for all rows $r_j$ with $j < i$, $\denote{\assume
          t.\matchrow(r_j)} = \emptyset$, and so the following holds: \[
          \denote{
          \begin{array}{l}
            \assume t.\matchrow(r_j);\\
            \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
            t.\action[\sigma(t)[i]](r_i.data)
        \end{array}}\,\pkt = \emptyset
          \]

        Now, for every index $j > i$, it may be that $\denote{\assume
          t.\matchrow(r_j)} \neq \emptyset$, but because $\pkt \models
        t.\matchrow(r_i)$, and $i < j$, we know $\denote{\bigwedge_{j=1}^{i-1}
          \neg t.\matchrow(r_j)} = \emptyset$, so conclude that \[
          \denote{
          \begin{array}{l}
            \assume t.\matchrow(r_j);\\
            \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
            t.\action[\sigma(t)[i]](r_i.data)
        \end{array}}\,\pkt = \emptyset
          \]

        Similarly, $\denote{\assume \bigwedge_{i=1}^{n} \neg t.\matchrow(r_j)} =
        \emptyset$, so conclude that $\denote{\assume \bigwedge_{i=1}^{n} \neg
          t.\matchrow(r_j);\SKIP} = \emptyset$

        Conclude that $\denote{t.\apply()[\sigma]} =
        \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. The result
        follows by applying the IH to $t.\action[r_i.\action](r_i.\data)$.

        \item[\textit{Case 2.}] Assume that for all $i \in \{1, \ldots, n\}$,
          $\pkt \not\models t.\matchrow(r_i)$. In this case,
          $\denote{t.\apply()}^\sigma\,\pkt = \{\pkt\}$. Show that $\denote{t.\apply()[\sigma]}\,\pkt = \{\}\pkt\}$.

          First observe that every $i \in \{1, \ldots, n\}$, we have
          $\denote{\assume t.\matchrow(r_i)}\,\pkt = \emptyset$, so conclude
          that
          \[ \forall i \in \{1,\ldots,n\}.\denote{
          \begin{array}{l}
            \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
            t.\action[\sigma(t)[i]](r_i.\data)
        \end{array}}\,\pkt = \emptyset
          \]

          Finally $\denote{\assume \bigwedge_{i=1}^n \neg
            t.\matchrow(r_i)}\,\pkt = \{\pkt\}$, and $\denote{\SKIP}\,\pkt
          =\{\pkt\}$, so conclude that $\denote{\assume \bigwedge_{i=1}^n \neg
            t.\matchrow(r_i);\SKIP}\,\pkt = \{\pkt\}$.
          Conclude that $\denote{t.\apply()[\sigma]}\pkt = \{pkt\}$.
      \end{enumerate}
  \end{enumerate}
\end{proof}

@@

\begin{figure}[H]
  \[
  \begin{array}{r >{\triangleq}c l}
    \WP(h.f := e, b)
    && b[e/h.f] \\
    \WP(\havoc x, b)
    && \forall x.\,b \\
    \WP(\assert b_a, b)
    && b_a \wedge b \\
    \WP(\assume b_a, b)
    && b_a \Rightarrow b \\
    \WP(t.\apply,b) && \textrm{\textcolor{red}{undefined}} \\
    \WP(c_1;c_2, b)
    && \WP(c_1,\WP(c_2,b)) \\
    \WP(c_1 \choice c_2, b)
    && \WP(c_1,b) \wedge \WP(c_2,b)
  \end{array}
  \]
  \caption{Dijkstra's weakest precondition function}
  \label{fig:wp}
\end{figure}

@@

\begin{theorem}[VC Correctness]
  Given a set of variables $V$, for every $c \in \Cmd$ such that $\fvs(c) = V$,
  and $\sigma \in \State$, the following formula is valid:
  \[ \WP(c[\sigma], \TRUE) \iff (\forall \vec{\rho_t}. \forall \vec{\gamma_t}. \state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\]
\end{theorem}

\begin{proof}
  Let $c$ and $\sigma$ be arbitary.
  Theorem~\ref{thm:instr-equiv} gives the equivalence
  \[\forall \pkt. \denote{c}^\sigma\,\pkt =_{\fvs(c)} \denote{\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr}\,\pkt\]

  By Theorem~\ref{thm:subst}, we get

  \[\forall \pkt.\denote{c[\sigma]}\,\pkt =_{\fvs(c)}  \denote{\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr}\,\pkt\]

  By Dijkstra '75, we can see

  \[\WP(c[\sigma], \TRUE) \iff  \WP(\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr, \TRUE))\]

  By definition of $\WP$, this is equivalent to

  \[\WP(c[\sigma],\TRUE) \iff \left(\forall \vec{\rho_t}.\forall \vec{\gamma_t}.\state\sigma_c \Rightarrow \WP(c\instr, \TRUE)\right)\]

\end{proof}

@@

\section{Abduction}
